Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-09-22 ms.Rnw
Compiled results from RL3

A brief analysis of the fitted parameters $\tilde{\paramVec}$ provides some interesting insight.
Consider the unexpectedly large exponent on rainfall: $r = \Sexpr{signif(p3s["r"],3)}$ (Eq.~\ref{eq:model3:water}).
Because the value of rainfall $J(t) \in (0, 1)$, the large coefficient results in $J(t)^r \approx 0$.
This result may initially be surprising given the importance of rainfall as a seasonal diver of cholera infections in Haiti (see Sec.~\ref{sec:science}).
One plausible explanation of this result is that the fitting procedure---used to obtain parameter values in the independent department model given the reported cases in other departments---has learned an over-dependence on the reported cholera cases in other departments, and thus the seasonality of cholera incidence in department $u$ depends on the highly correlated covariate of the reported cholera cases in departments $1:U\setminus \{u\}$.
Whatever the case may be, this fitting procedure results in model likelihoods unable to improve upon a log-linear benchmark (see Sec.~\ref{sec:model_diagnostics} and Tab.~\ref{tab:likes}), and highlights the need for advancements in statistical methodology that permit inference on models with coupled metapopulation dynamics.

@@

\begin{figure}[ht]
<<Plot_Model1_Scenarios, fig.height=2>>=
lay <- rbind(c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5))
gridExtra::grid.arrange(
  gg_m1_V0, gg_m1_V1, gg_m1_V2, gg_m1_V3, gg_m1_V4,
  layout_matrix = lay
)
@
\caption{\label{fig:Mod1Scenarios}
Simulations of Model~1 under each vaccination scenario.
Blue line indicates the median of model simulations, and ribbon represents $95\%$ of simulations.
The various vaccination campaigns made no practical difference in the median scenario, but a drastic difference in the extreme cases.
}
\end{figure}

@@

\begin{figure}[ht]
<<Plot_Model3_Scenarios, fig.height=2>>=
lay <- rbind(c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5))
gridExtra::grid.arrange(
  gg_m3_V0, gg_m3_V1, gg_m3_V2, gg_m3_V3, gg_m3_V4,
  layout_matrix = lay
)
@
\caption{\label{fig:Mod3Scenarios}
Simulations of Model~3 under each vaccination scenario.
Blue line indicates the median of model simulations, and ribbon represents $95\%$ of simulations.
The various vaccination campaigns made no practical difference in the median scenario, but a drastic difference in the extreme cases.
}
\end{figure}

>>>>

A brief analysis of the fitted parameters $\tilde{\paramVec}$ provides some interesting insight.
Consider the large estimated exponent on rainfall: $r = \Sexpr{signif(p3s["r"],3)}$ (Eq.~\ref{eq:model3:water}).
Because the value of rainfall $J(t) \in (0, 1)$, the large coefficient results in $J(t)^r \approx 0$.
This result may initially be surprising given the importance of rainfall as a seasonal diver of cholera infections in Haiti (see Sec.~\ref{sec:science}).
One plausible explanation of this result is that the fitting procedure---used to obtain parameter values in the independent department model given the reported cases in other departments---has learned an over-dependence on the reported cholera cases in other departments, and thus the seasonality of cholera incidence in department $u$ depends on the highly correlated covariate of the reported cholera cases in departments $1:U\setminus \{u\}$.
Whatever the case may be, this fitting procedure results in a model that is unable to replicate the observed number of cases by simulating from initial conditions (see Fig.~\ref{h3spatsims}), and highlights the need for advancements in statistical methodology that permit inference on models with coupled metapopulation dynamics.

@@

\begin{figure}[ht]
<<Plot_Model1_Scenarios, fig.height=2>>=
lay <- rbind(c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5))
gridExtra::grid.arrange(
  gg_m1_V0, gg_m1_V1, gg_m1_V2, gg_m1_V3, gg_m1_V4,
  layout_matrix = lay
)
@
\caption{\label{fig:Mod1Scenarios}
Simulations of Model~1 under each vaccination scenario.
Blue line indicates the simulated median of reported cases, and the ribbon represents $95\%$ of simulations.
% The various vaccination campaigns made no practical difference in the median scenario, but a drastic difference in the extreme cases.
}
\end{figure}

@@

\begin{figure}[ht]
<<Plot_Model3_Scenarios, fig.height=2>>=
lay <- rbind(c(1, 1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4, 5, 5, 5))
gridExtra::grid.arrange(
  gg_m3_V0, gg_m3_V1, gg_m3_V2, gg_m3_V3, gg_m3_V4,
  layout_matrix = lay
)
@
\caption{\label{fig:Mod3Scenarios}
Simulations of Model~3 under each vaccination scenario.
Blue line indicates the simulated median of reported cases, and ribbon represents $95\%$ of simulations.
% The various vaccination campaigns made no practical difference in the median scenario, but a drastic difference in the extreme cases.
}
\end{figure}
